Multi-scale trend analysis of water quality using error propagation of generalized additive models

### 2020-11-13 manu_draft.Rmd
code chunk updated to wqtrends one model in manu draft

thm <-theme_minimal() + 
  theme(
    legend.title = element_blank(),
    legend.position = 'right'
  )
lncol <- 'tomato1'
alph <- 0.5

p1 <- show_prddoy(mods = mods, ylab = 'Chlorophyll-a (ug/L)', nfac = 3, size = 0.8, alpha = 0.8) + 
  guides(colour = ggplot2::guide_colourbar(barheight = 15, barwidth = 1)) +
  scale_colour_gradient(low = "white", high = "black") + 
  thm + 
  labs(
    title = '(a) Model predictions by day of year'
  )

prds <- anlz_prd(mods = mods, trans = 'log10')

@@

      mod <- get(nm) %>% 
        pull(model)

      prd <- anlz_prd(mods = mod) %>% 
        anlz_backtrans()

@@

```{r, results = 'hide', message = F}
toplo <- modslog_chl34 %>%
  pull(model)

chg1 <- anlz_perchg(mods = toplo, baseyr = 1991, testy = 2000)
chg2 <- anlz_perchg(mods = toplo, baseyr = 2000, testy = 2010)
chg3 <- anlz_perchg(mods = toplo, baseyr = 2010, testy = 2019)
chg <- bind_rows(chg1, chg2, chg3) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  mutate(
    pval = p_ast(pval)
  )

@@

      out <- anlz_avgseason(mods = toplo, doystr = doystr, doyend = doyend) %>% 
        anlz_mixmeta(yrstr = yrstr, yrend = yrend) %>% 
        .[[1]] %>% 
        summary %>% 
        coef(.) %>% 
        .[2, ] 
      out <- data.frame(est = out[['Estimate']], pval = out[['Pr(>|z|)']])

@@

p1 <- show_perchg(mods = toplo, baseyr = 1991, testyr = 2000, ylab = ylab) +
  labs(
    subtitle = paste0('(a) 1991: ', chg[[1, 2]],' ug/L, 2000: ', chg[[1, 3]], 'ug/L\nChange: ', chg[[1,4]], '%, ', chg[[1,5]]), 
    title = NULL
  )
p2 <- show_perchg(mods = toplo, baseyr = 2000, testyr = 2010, ylab = NULL) +
  labs(
    subtitle = paste0('(a) 2000: ', chg[[2, 2]],' ug/L, 2010: ', chg[[2, 3]], 'ug/L\nChange: ', chg[[2, 4]], '%, ', chg[[2, 5]]), 
    title = NULL
  )
p3 <- show_perchg(mods = toplo, baseyr = 2010, testyr = 2019, ylab = NULL) +
  labs(
    subtitle = paste0('(a) 2010: ', chg[[3, 2]],' ug/L, 2019: ', chg[[3, 3]], 'ug/L\nChange: ', chg[[3, 4]], '%, ', chg[[3, 5]]), 
    title = NULL
  )
p4 <- show_avgseason(mods = toplo, doystr = 41, doyend = 213, yrstr = 1991, yrend = 2000, ylab = ylab) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[1]],
    title = NULL
  )
p5 <- show_avgseason(mods = toplo, doystr = 41, doyend = 213, yrstr = 2000, yrend = 2010, ylab = NULL) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[2]], 
    title = NULL
  )
p6 <- show_avgseason(mods = toplo, doystr = 41, doyend = 213, yrstr = 2010, yrend = 2019, ylab = NULL) +
  coord_cartesian(ylim = ylim) +  
  labs(
    subtitle = seas[[3]], 
    title = NULL
  )
p7 <- show_avgseason(mods = toplo, doystr = 213, doyend = 338, yrstr = 1991, yrend = 2000, ylab = ylab) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[4]], 
    title = NULL
  )
p8 <- show_avgseason(mods = toplo, doystr = 213, doyend = 338, yrstr = 2000, yrend = 2010, ylab = NULL) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[5]], 
    title = NULL
  )
p9 <- show_avgseason(mods = toplo, doystr = 213, doyend = 338, yrstr = 2010, yrend = 2019, ylab = NULL) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[6]],
    title = NULL
  )

@@

toplo <- seastrnd3 %>% 
  na.omit() %>% 
  mutate(
    sta = gsub('^data/modslog_chl|\\.RData', '', fl), 
    subttl = ifelse(doystr < 150, '(a) Jan-Jun', '(b) Jul-Dec'),
    pval = dplyr::case_when(
      pval < 0.05 ~ 'p < 0.05', 
      T ~ 'ns'
    ), 
    pval = factor(pval, levels = c('ns', 'p < 0.05'))
  ) %>% 
  dplyr::filter(yr <= 2016 & yr >= 1993)

p <- ggplot(data = toplo, aes(x = yr, y = yrcoef, fill = pval)) + 
  geom_point(shape = 21, size = 3) +
  scale_x_continuous(limits = c(1993, 2016), breaks = seq(1993, 2016)) +
  scale_y_continuous(limits = c(-2, 2)) +
  scale_fill_manual(values = c(NA, 'tomato1'), drop = FALSE) +
  geom_hline(yintercept = 0) + 
  facet_grid(sta ~ subttl) + 
  theme_minimal(base_family = 'serif') + 
  theme(
    axis.title.x = element_blank(), 
    legend.position = 'right', 
    legend.title = element_blank(),
    strip.background = element_blank(), 
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(), 
    panel.grid.minor.y = element_blank(), 
    panel.grid.major.y = element_blank(), 
    strip.text.x = element_text(size = 12, hjust = 0), 
    axis.ticks = element_line(), 
    panel.spacing.y = grid::unit(1, "lines")
  ) +
  labs(
    y = expression(paste('Chlorophyll change ( ', mu, 'g ', L^-1, yr^-1, ')'))
  )

>>>>

thm <-theme_minimal() + 
  theme(
    legend.title = element_blank(),
    legend.position = 'right'
  )
lncol <- 'tomato1'
alph <- 0.5
ylm <- c(0.8, 100)

p1a <- show_prddoy(mods[[1]], ylab = 'SY: Chl-a (ug/L)', size = 0.8, alpha = 0.8) + 
  guides(colour = ggplot2::guide_colourbar(barheight = 15, barwidth = 1)) +
  scale_colour_gradient(low = "white", high = "black") + 
  coord_cartesian(ylim = ylm) +
  theme_minimal() + 
  theme(
    legend.position = 'none', 
    axis.title.x = element_blank()
  ) +
  labs(
    title = '(a) Model predictions by day of year'
  )
p1b <- show_prddoy(modstr[[2]], ylab = 'SYD: Chl-a (ug/L)', size = 0.8, alpha = 0.8) + 
  guides(colour = ggplot2::guide_colourbar(barheight = 15, barwidth = 1)) +
  scale_colour_gradient(low = "white", high = "black") + 
  coord_cartesian(ylim = ylm) +
  theme_minimal() + 
  theme(
    legend.position = 'none'
  )
p1c <- show_prddoy(mods[[3]], ylab = 'SYDI: Chl-a (ug/L)', size = 0.8, alpha = 0.8) + 
  guides(colour = ggplot2::guide_colourbar(barheight = 15, barwidth = 1)) +
  scale_colour_gradient(low = "white", high = "black") + 
  coord_cartesian(ylim = ylm) +
  theme_minimal() + 
  theme(
    legend.position = 'right', 
    axis.title.x = element_blank()
  )

p1 <- p1a + p1b + p1c + plot_layout(ncol = 3, widths = c(1, 1, 1))
prds <- lapply(modstr, anlz_prd) %>% 
  enframe('model', 'dat') %>%
  unnest('dat')

@@

      mod <- get(nm) %>% 
        pull(model) %>% 
        deframe()

      prd <- anlz_prd(mod) %>% 
        anlz_backtrans()

@@

```{r, results = 'hide', message = F}
mod <- modslog_chl34 %>%
  pull(model) %>% 
  deframe()

chg1 <- anlz_perchg(mod, baseyr = 1991, testy = 2000)
chg2 <- anlz_perchg(mod, baseyr = 2000, testy = 2010)
chg3 <- anlz_perchg(mod, baseyr = 2010, testy = 2019)
chg <- bind_rows(chg1, chg2, chg3) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  mutate(
    pval = p_ast(pval)
  )

@@

      out <- anlz_avgseason(mod, doystr = doystr, doyend = doyend) %>% 
        anlz_mixmeta(yrstr = yrstr, yrend = yrend) %>% 
        summary %>% 
        coef(.) %>% 
        .[2, ] 
      out <- data.frame(est = out[['Estimate']], pval = out[['Pr(>|z|)']])

@@

p1 <- show_perchg(mod, baseyr = 1991, testyr = 2000, ylab = ylab) +
  labs(
    subtitle = paste0('(a) 1991: ', chg[[1, 1]],' ug/L, 2000: ', chg[[1, 2]], 'ug/L\nChange: ', chg[[1,3]], '%, ', chg[[1,4]]), 
    title = NULL
  )
p2 <- show_perchg(mod, baseyr = 2000, testyr = 2010, ylab = NULL) +
  labs(
    subtitle = paste0('(a) 2000: ', chg[[2, 1]],' ug/L, 2010: ', chg[[2, 2]], 'ug/L\nChange: ', chg[[2, 3]], '%, ', chg[[2, 4]]), 
    title = NULL
  )
p3 <- show_perchg(mod, baseyr = 2010, testyr = 2019, ylab = NULL) +
  labs(
    subtitle = paste0('(a) 2010: ', chg[[3, 1]],' ug/L, 2019: ', chg[[3,2]], 'ug/L\nChange: ', chg[[3, 3]], '%, ', chg[[3, 4]]), 
    title = NULL
  )
p4 <- show_avgseason(mod, doystr = 41, doyend = 213, yrstr = 1991, yrend = 2000, ylab = ylab) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[1]],
    title = NULL
  )
p5 <- show_avgseason(mod, doystr = 41, doyend = 213, yrstr = 2000, yrend = 2010, ylab = NULL) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[2]], 
    title = NULL
  )
p6 <- show_avgseason(mod, doystr = 41, doyend = 213, yrstr = 2010, yrend = 2019, ylab = NULL) +
  coord_cartesian(ylim = ylim) +  
  labs(
    subtitle = seas[[3]], 
    title = NULL
  )
p7 <- show_avgseason(mod, doystr = 213, doyend = 338, yrstr = 1991, yrend = 2000, ylab = ylab) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[4]], 
    title = NULL
  )
p8 <- show_avgseason(mod, doystr = 213, doyend = 338, yrstr = 2000, yrend = 2010, ylab = NULL) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[5]], 
    title = NULL
  )
p9 <- show_avgseason(mod, doystr = 213, doyend = 338, yrstr = 2010, yrend = 2019, ylab = NULL) +
  coord_cartesian(ylim = ylim) + 
  labs(
    subtitle = seas[[6]],
    title = NULL
  )

@@

toplo <- seastrnd3 %>% 
  na.omit() %>% 
  mutate(
    subttl = ifelse(doystr < 150, '(a) Jan-Jun', '(b) Jul-Dec'),
    pval = dplyr::case_when(
      pval < 0.05 ~ 'p < 0.05', 
      T ~ 'ns'
    ), 
    pval = factor(pval, levels = c('ns', 'p < 0.05'))
  ) %>% 
  dplyr::filter(yr <= 2016 & yr >= 1993)

p <- ggplot(data = toplo, aes(x = yr, y = yrcoef, fill = pval)) + 
  geom_point(shape = 21, size = 3) +
  scale_x_continuous(limits = c(1993, 2016), breaks = seq(1993, 2016)) +
  scale_y_continuous(limits = c(-2, 2)) +
  scale_fill_manual(values = c(NA, 'tomato1'), drop = FALSE) +
  geom_hline(yintercept = 0) + 
  facet_grid(station ~ subttl) + 
  theme_minimal(base_family = 'serif') + 
  theme(
    axis.title.x = element_blank(), 
    legend.position = 'right', 
    legend.title = element_blank(),
    strip.background = element_blank(), 
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(), 
    panel.grid.minor.y = element_blank(), 
    panel.grid.major.y = element_blank(), 
    strip.text.x = element_text(size = 12, hjust = 0), 
    axis.ticks = element_line(), 
    panel.spacing.y = grid::unit(1, "lines")
  ) +
  labs(
    y = expression(paste('Chlorophyll change ( ', mu, 'g ', L^-1, yr^-1, ')'))
  )
